On the number of tiles touched by a\\ line segment on a rectangular grid

### 2021-08-02 paper.tex
More writing on the average number

% **Aq. Prepare and include two subfigures
Given the coordinates $(x_1, y_1)$ of the first segment endpoint, with $0 \leq x_1 \leq a$, $0 \leq y_1 \leq b$, the second endpoint $(x_2, y_2)$ lies on a circle with radius $\len$ centered at $(x_1,y_1)$, as shown in Figure **ab**. The segment orientation is a random angle $\orient$ uniformly distributed on $[0,2\pi)$. It is easy to see that $i \geq n$ if and only if either $x_2 \geq a(n-1)$ or $x_2 \leq -a(n-2)$. The latter two events are exclusive, and by symmetry have the same probability. Therefore
\begin{equation}
\label{eq: Pr i = 2 Pr}
\Pr[i \geq n] = 2 \Pr[x_2 \geq a(n-1)].
\end{equation}

@@

\begin{theorem}
\label{theo: funta, form}
Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above. The average number of tiles touched by the segment, $\funta(\len)$, is 
\begin{equation}
\label{eq: funta, expression}
\funta(\len) = \frac{2\len}{\pi}\left(\frac 1 a + \frac 1 b\right) + 1.
\end{equation}
\end{theorem}

\begin{proof}
***
\end{proof}

>>>>

% *** Prepare and include two subfigures
Given the coordinates $(x_1, y_1)$ of the first segment endpoint, with $0 \leq x_1 \leq a$, $0 \leq y_1 \leq b$, the second endpoint $(x_2, y_2)$ lies on a circle with radius $\len$ centered at $(x_1,y_1)$, as shown in Figure **ab**. The segment orientation is a random angle $\orient$ uniformly distributed on $[0,2\pi)$. It is easy to see that $i \geq n$ if and only if either $x_2 \geq a(n-1)$ or $x_2 \leq -a(n-2)$. The latter two events are exclusive, and by symmetry have the same probability. Therefore
\begin{equation}
\label{eq: Pr i = 2 Pr}
\Pr[i \geq n] = 2 \Pr[x_2 \geq a(n-1)].
\end{equation}

@@

\begin{theorem}
\label{theo: funta, form}
Given $a, b, \len \in \mathbb R^+$, consider a grid with parameters $a, b$ and a uniformly random segment of length $\len$, as defined above. The average number of tiles touched by the segment, $\funta(\len)$, is 
\begin{equation}
\label{eq: funta, form}
\funta(\len) = \frac{2\len}{\pi}\left(\frac 1 a + \frac 1 b\right) + 1.
\end{equation}
\end{theorem}

\begin{proof}
The expected value of $i$ can be obtained as
\begin{equation}
\label{eq: theo funta form 1}
\E[i] = \sum_{n=1}^\infty \Pr[i \geq n].
\end{equation}
From \eqref{eq: Pr i geq n}, this sum only has a finite number of non-zero terms: $1$ for $n=1$, plus $\lceil \len/a \rceil-1$ terms given by the second line of \eqref{eq: Pr i geq n}, plus one term given by the third line, corresponding to $n = \lceil \len/a \rceil +1 $:
\begin{equation}
\label{eq: theo funta form 2}
\begin{split}
\E[i] = 1 +
\sum_{n=2}^{\lceil \len/a \rceil} \frac{2\len}{\pi a} \left(g\left(\frac{a(n-1)}{\len}\right)-g\left(\frac{a(n-2)}{\len} \right)\right) + \\
\frac{2\len}{\pi a} \left(1-g\left(\frac{a(\lceil \len/a \rceil - 1)}{\len} \right)\right).
\end{split}
\end{equation}
Most terms involving the function $g$ in \eqref{eq: theo funta form 2} cancel, with the result
\begin{equation}
\label{eq: theo funta form 3}
\E[i] = 1+\frac{2\len}{\pi a} (1-g(0))= 1+\frac{2\len}{\pi a}.
\end{equation}
Similarly,
\begin{equation}
\label{eq: theo funta form 4}
\E[j] = 1+\frac{2\len}{\pi b}.
\end{equation}
Substituting \eqref{eq: theo funta form 3} and \eqref{eq: theo funta form 4} into \eqref{eq: funta E[i] E[j]} gives \eqref{eq: funta, form}.
\end{proof}
